Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-06-06 manuscript/revision2/revision2.tex
revision2: Add SI Appendix Sections on BLUPs, GWAS, SNP subset selection for mash

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}

\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (22). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or alleles with effects with opposing fitness
outcomes (23--26), previous empirical work has found limited evidence of
antagonistic pleiotropy (12, 27). However, this work has been limited by
a known statistical bias that reduced detection of genetic effects that
differ in sign (27--29). Thus, despite substantial interest in the
frequencies of various forms of GxE, particularly sign-changing GxE,
relative to other forms of GxE remains unknown.

@@

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology for cloned
individuals grown at eight common gardens to many patterns of weather
covariance at these gardens. To do this, we phenotype a diversity panel
of hundreds of switchgrass genotypes from the Midwest and Gulf
subpopulations for the timing of vegetative and reproductive development
at eight common garden locations spanning 17 degrees of latitude. These
gardens cover the majority of the latitudinal and climatic range of
switchgrass and capture the most comprehensive picture to date of the
environmental variation this species encounters. We define multiple ways
phenological timings might covary with weather (Table 1) and additional
ways phenological timings might vary by garden (SI Appendix, Section
S1), then jointly re-estimate genetic effects on these timings at all
eight common gardens using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included (SI
Appendix, Section S2) (36). We use the Bayesian framework \emph{mash}
(multivariate adaptive shrinkage) developed by (36), to refine effect
size estimates from genome-wide association (GWAS) conducted on
individuals at each of the eight gardens. \emph{mash} allows us to
identify and specify multiple covariance structures among genetic effect
estimates across sites, including structures that represent covariance
in weather variables of interest. Importantly, this method circumvents
statistical biases in detecting genetic effects with the same or
opposite signs (37). To confirm our genetic mapping of GxWeather, we
compare our significant posterior genetic effects locations from
\emph{mash} to mapping results from an outbred mapping population grown
at the same sites. Our analyses allow us to describe the weather cues
and genetic variation affecting phenology in two divergent natural
populations of switchgrass across the species' latitudinal range.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects of SNPs across all eight common gardens using the
\emph{mash} model, which can flexibly capture many modes of covariance
across gardens (Figure~\ref{fig-map} D; SI Appendix, Datasets 1-6). We
specified three qualitative categories of covariance structures across
gardens: ``canonical'' covariance, with simple patterns of effect size
covariance introduced in the initial \emph{mash} manuscript;
``data-driven'' covariances derived from common patterns of SNP effects
observed in the data, and ``GxWeather'' covariance, estimated from the
covariance of empirical weather patterns at each garden at specific
times before the phenological event (Table 1; SI Appendix, Section S2).
For example, the more similar that the amount of rainfall prior to
flowering is for genotypes A and B between gardens in MI and MO, the
nearer these covariances will be to one. If rainfall amounts are
uncorrelated, covariance will be near zero; if there is rank changing
(e.g.~high rainfall before flowering of A at MO but not MI, and high
rainfall before flowering for B at MI but not MO), this covariance will
be near negative one. We specified multiple covariance structures for
the canonical, data-driven, and GxWeather categories, each of which
represented a specific pattern of covariance in genetic effect size
estimates that could be included in the model. The phenotypic
correlations for the timing of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the \emph{mash} model.

@@

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited a difference in effect sign, between pairs of
Texas and North gardens (Figure~\ref{fig-effects} B; SI Appendix, Fig
SX). 78.7\% of pairwise comparisons between North and Texas gardens had
a difference in sign, while only 28.6\% and 0.2\% of North-North or
Texas-Texas comparisons had a difference in sign, respectively
(Figure~\ref{fig-effects} B). The majority of pairwise effects for
greenup for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
subpopulations were the same sign, and effects most often differed in
magnitude between gardens within and between the regions
(Figure~\ref{fig-effects} B).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited a difference in effect sign within or between regions
(Figure~\ref{fig-effects} B). More effects differed in magnitude between
the Texas and North regions than within these regions (42.7\% vs
\textless20\%; Figure~\ref{fig-effects} B), while the majority of
effects had no GxE. The Midwest population had relatively few
significant effects for flowering, but a large proportion of these
differed in sign between Texas and North regions (42.7\%) or within the
North region (65.4\%). Finally, in Both subpopulations, less than 20\%
of pairwise effects differed in sign (Figure~\ref{fig-effects} B). Most
differences in sign were between TX1, the southernmost garden, and all
other gardens (SI Appendix, Fig SX). Similarly, more effects that
differed in magnitude included gardens in the Texas region
(52.3-55.9\%), and most effect pairs in the North region were not
distinguishable (91.5\%).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7). All flowering and
green-up QTL intervals contained at least one SNP significant in at
least one \emph{mash} run at a log10-transformed Bayes Factor
\textgreater{} 2, or in the 1\% tail of significance, whichever was
stricter (SI Appendix, Dataset 8). We also looked for enrichments of
\emph{mash} SNPs in the 1\% tail of significance (the `\emph{mash} 1\%
tail') within each QTL interval. At the 5\% level, three QTL had
enrichments of SNPs in the \emph{mash} 1\% tail. Overall, there were
five significant enrichments (p \textless{} 0.05, hypergeometric test)
of SNPs in the \emph{mash} 1\% tail in the QTL intervals. Thus, we were
able to experimentally support some of our re-estimates of SNP effects
with a QTL mapping experiment using a separate mapping population.

@@

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated effects in eight common gardens, for effects with
lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
sites with three types of GxE. Sign: Effects that differ in sign at
these pairs of gardens (p \textless{} 0.05, lfsr). Magnitude: Effects
identical in sign (p \textless{} 0.05, lfsr) that differ in magnitude by
a factor of \textgreater0.4. None: Effects not distinguishable by
magnitude nor sign of the effect. b) The fraction of effects with each
GxE type for the start of vegetative growth (Greenup Date) and
reproductive growth (Flowering Date), within and between two genetic
subpopulations. Common gardens are grouped by the larger region they
came from: North gardens are within the natural range of the Midwest
subpopulation, while Texas gardens are within the natural range of the
Gulf subpopulation.}

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (35, 39, 40). We scored the
start of vegetative growth, or green-up date, as the day of the year
when 50\% of the tiller area of the crown of the plant cut the previous
year had green growth. The start of reproductive growth, or flowering
date, was the day of the year when 50\% of the plant tillers had
panicles undergoing anthesis. We scored green-up and flowering as day of
the year, then linked these dates to multiple weather-based
environmental factors measured daily at each common garden (SI Appendix,
Section S1, Table S1).

The formation and resequencing of the diversity panel has been described
previously (33). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (41), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both' subpopulations). Analyses to determine
narrow-sense heritability (h\textsuperscript{2}) for green-up and
flowering were done using linear mixed models and followed (33). Details
on these models can be found in (SI Appendix, Section S2).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(36). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (40). We used the effect estimates for single
nucleotide polymorphisms (SNPs) from these GWAS as preliminary effect
estimates for the two traits at each garden. We then jointly modeled
genetic effect estimates, and the mixture of ways these effects might
covary, across all eight common gardens, using a two-part procedure. We
first modeled genetic effects with different combinations of covariance
structures, using the effect size estimates from a subset of \ul{19K}
relatively unlinked (r\textsuperscript{2} \textless{} 0.2), randomly
selected SNPs used in the GWAS. In these models (hereafter referred to
as ``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Details on these models can be found in (SI
Appendix, Section S3). Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2) SNPs from each univariate GWAS that had the largest
effect estimates in each GWAS (hereafter ``strong'' effects), which gave
sets of \ul{\textbf{48K - 80K}} SNPs, and jointly re-estimated these
effect size estimates at all eight sites using the set of significant
covariance matrices that most improved the model fit on the random
effects. Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS.

We generated hypothesis-based covariance matrices derived from
correlations in environmental cues in the green-up or flowering date
windows for the three populations (Table 1; SI Appendix, Table S1,
Section S1). These covariance matrices represent correlations between
identical genotypes drawn from a specific population at pairs of common
gardens; covariances near one mean that the population has a strong,
positive linear relationship in individual responses at that pair of
gardens, while covariances near zero mean that there is no relationship
within the population for individual responses at that pair of gardens.
\emph{Mash} SNP effects will undergo strong shrinkage towards one
another in the first case, and little shrinkage in the second case.
\emph{Mash} also generates data-driven covariance matrices corresponding
to major patterns of SNP effects present in the data. We generated six
data-driven matrices per \emph{mash} run, five produced by singular
value decomposition (SVD) of an overall matrix.

@@

To confirm candidate genomic regions and patterns of allelic effects
found in the diversity panel, we analyzed flowering in an outbred
pseudo-F2 cross between four individuals, two Midwest and two Gulf
individuals. The formation of this mapping population has been described
previously (35); additional details on QTL mapping can be found in SI
Appendix, Section S4. To be directly comparable to the diversity panel
data, only 2019 phenology data from the pseudo-F2 cross from the same
eight common garden sites were used.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

Flowering time, or the transition from vegetative to reproductive
growth, is a common subject of GxE research (5--11), a key output of
selection driving adaptation to local environments (2, 12, 13), and a
selection target for crop improvement to adapt crops to local or future
environments (14). Changing flowering responsiveness to photoperiod cues
has allowed geographic range expansion and increased yields in several
cereal species (15--19) and other crops (20, 21). Recent statistical
advances in studying phenological GxE have involved determining critical
environmental indices before the phenological event occurs, such as
photothermal time within a critical growth window (10). However, most
studies of flowering GxE focus on finding a single, best fitting form of
genotype-environment covariance, despite the key expectation that
different genetic subpopulations, and even different genomic regions,
have likely evolved distinct patterns of GxE (22). Additionally, despite
theoretical predictions that local adaptation should involve
antagonistic pleiotropy, or alleles with effects with opposing fitness
outcomes (23--26), previous empirical work has found limited evidence of
antagonistic pleiotropy (12, 27). However, this work has been limited by
a known statistical bias that reduced detection of genetic effects that
differ in sign (27--29). Thus, despite substantial interest in the
frequencies of various forms of GxE, the frequency of sign-changing GxE
relative to other forms of GxE remains unknown.

@@

Here, we estimate GxWeather for two phenological timings in switchgrass
by assigning patterns of genetic effects on phenology for cloned
individuals grown at eight common gardens to many patterns of weather
covariance at these gardens. To do this, we phenotype a diversity panel
of hundreds of switchgrass genotypes from the Midwest and Gulf
subpopulations for the timing of vegetative and reproductive development
at eight common garden locations spanning 17 degrees of latitude. These
gardens cover the majority of the latitudinal and climatic range of
switchgrass and capture the most comprehensive picture to date of the
environmental variation this species encounters. We define multiple ways
phenological timings might covary with weather (Table 1) and additional
ways phenological timings might vary by garden (SI Appendix, Section
S1), then jointly re-estimate genetic effects on these timings at all
eight common gardens using the set of these covariance matrices that
significantly improved the modeled log-likelihood when included (SI
Appendix, Section S2-S5) (36). We use the Bayesian framework \emph{mash}
(multivariate adaptive shrinkage) developed by (36), to refine effect
size estimates from genome-wide association (GWAS) conducted on
individuals at each of the eight gardens. \emph{mash} allows us to
identify and specify multiple covariance structures among genetic effect
estimates across sites, including structures that represent covariance
in weather variables of interest. Importantly, this method circumvents
statistical biases in detecting genetic effects with the same or
opposite signs (37). To confirm our genetic mapping of GxWeather, we
compare our significant posterior genetic effects locations from
\emph{mash} to mapping results from an outbred mapping population grown
at the same sites. Our analyses allow us to describe the weather cues
and genetic variation affecting phenology in two divergent natural
populations of switchgrass across the species' latitudinal range.

@@

We next looked for evidence of GxWeather by jointly re-estimating
genetic effects of SNPs across all eight common gardens using the
\emph{mash} model, which can flexibly capture many modes of covariance
across gardens (Figure~\ref{fig-map} D; SI Appendix, Sections S1-S5,
Datasets 1-6). We specified three qualitative categories of covariance
structures across gardens: ``canonical'' covariance, with simple
patterns of effect size covariance introduced in the initial \emph{mash}
manuscript; ``data-driven'' covariances derived from common patterns of
SNP effects observed in the data, and ``GxWeather'' covariance,
estimated from the covariance of empirical weather patterns at each
garden at specific times before the phenological event (Table 1; SI
Appendix, Section S1). For example, the more similar that the amount of
rainfall prior to flowering is for genotypes A and B between gardens in
MI and MO, the nearer these covariances will be to one. If rainfall
amounts are uncorrelated, covariance will be near zero; if there is rank
changing (e.g.~high rainfall before flowering of A at MO but not MI, and
high rainfall before flowering for B at MI but not MO), this covariance
will be near negative one. We specified multiple covariance structures
for the canonical, data-driven, and GxWeather categories, each of which
represented a specific pattern of covariance in genetic effect size
estimates that could be included in the model. The phenotypic
correlations for the timing of vegetative growth had moderate negative
correlations between the Texas and North gardens, particularly in the
Gulf subpopulation and Both subpopulations (Figure~\ref{fig-map} B). If
these phenotypic correlations have a genetic component, they could be
partially or completely captured by the covariance structures specified
in the \emph{mash} model.

@@

For green-up date for the Gulf subpopulation, hundreds to thousands of
pairwise effects exhibited a difference in effect sign, between pairs of
Texas and North gardens (Figure~\ref{fig-effects} B; SI Appendix, Fig
S1). 78.7\% of pairwise comparisons between North and Texas gardens had
a difference in sign, while only 28.6\% and 0.2\% of North-North or
Texas-Texas comparisons had a difference in sign, respectively
(Figure~\ref{fig-effects} B). The majority of pairwise effects for
greenup for the Midwest (\textgreater55\%) and Both (\textgreater85\%)
subpopulations were the same sign, and effects most often differed in
magnitude between gardens within and between the regions
(Figure~\ref{fig-effects} B).

For flowering date for the Gulf subpopulation, less than 2\% of pairwise
effects exhibited a difference in effect sign within or between regions
(Figure~\ref{fig-effects} B). More effects differed in magnitude between
the Texas and North regions than within these regions (42.7\% vs
\textless20\%; Figure~\ref{fig-effects} B), while the majority of
effects had no GxE. The Midwest population had relatively few
significant effects for flowering, but a large proportion of these
differed in sign between Texas and North regions (42.7\%) or within the
North region (65.4\%). Finally, in Both subpopulations, less than 20\%
of pairwise effects differed in sign (Figure~\ref{fig-effects} B). Most
differences in sign were between TX1, the southernmost garden, and all
other gardens (SI Appendix, Figure S1). Similarly, more effects that
differed in magnitude included gardens in the Texas region
(52.3-55.9\%), and most effect pairs in the North region were not
distinguishable (91.5\%).

@@

We sought additional experimental support the occurrence and location of
our significant SNP effect re-estimates using an independent pseudo-F2
mapping population created from Gulf \& Midwest individuals and grown at
the same sites (Figure~\ref{fig-qtl} A,B). We conducted quantitative
trait loci (QTL) mapping of flowering as functions of five environmental
cues that we also used as covariance matrices in \emph{mash}, and
identified eight QTL for flowering date, eight QTL for flowering as a
function of day length change two days prior, one QTL for the start of
vegetative growth, and two QTL for vegetative growth as a function of
daylength change one day prior, all of which showed QTL by environment
interactions (SI Appendix, Figure S2). All QTL for flowering overlapped
one or more homologs from rice or \emph{A. thaliana} with functionally
validated roles in flowering (SI Appendix, Dataset 7). All flowering and
green-up QTL intervals contained at least one SNP significant in at
least one \emph{mash} run at a log10-transformed Bayes Factor
\textgreater{} 2, or in the 1\% tail of significance, whichever was
stricter (SI Appendix, Dataset 8). We also looked for enrichments of
\emph{mash} SNPs in the 1\% tail of significance (the `\emph{mash} 1\%
tail') within each QTL interval. At the 5\% level, three QTL had
enrichments of SNPs in the \emph{mash} 1\% tail. Overall, there were
five significant enrichments (p \textless{} 0.05, hypergeometric test)
of SNPs in the \emph{mash} 1\% tail in the QTL intervals. Thus, we were
able to experimentally support some of our re-estimates of SNP effects
with a QTL mapping experiment using a separate mapping population.

@@

\caption{\label{fig-effects}Types of GxE present between of pairs of
jointly re-estimated effects in eight common gardens, for effects with
lfsr \textless{} 0.05. a) Examples of effect patterns at three pairs of
sites with three types of GxE. All effects are for an alternate allele,
with the reference allele effect defined to be zero at both gardens.
Sign: Effects that differ in sign at these pairs of gardens (p
\textless{} 0.05, lfsr). Magnitude: Effects identical in sign (p
\textless{} 0.05, lfsr) that differ in magnitude by a factor of
\textgreater0.4. None: Effects not distinguishable by magnitude nor sign
of the effect. b) The fraction of effects with each GxE type for the
start of vegetative growth (Greenup Date) and reproductive growth
(Flowering Date), within and between two genetic subpopulations. Common
gardens are grouped by the larger region they came from: North gardens
are within the natural range of the Midwest subpopulation, while Texas
gardens are within the natural range of the Gulf subpopulation.}

@@

In 2019, we scored two phenological events every two days in two mapping
populations of switchgrass, a diversity panel and a pseudo-F2 cross,
planted at eight common garden locations (35, 39, 40). We scored the
start of vegetative growth, or green-up date, as the day of the year
when 50\% of the tiller area of the crown of the plant cut the previous
year had green growth. The start of reproductive growth, or flowering
date, was the day of the year when 50\% of the plant tillers had
panicles undergoing anthesis. We scored green-up and flowering as day of
the year, then linked these dates to multiple weather-based
environmental factors measured daily at each common garden (Table 1, SI
Appendix, Section S1).

The formation and resequencing of the diversity panel has been described
previously (33). The diversity panel contained 134 sequenced, clonally
propagated individuals from the Midwest genetic subpopulation, and 229
from the Gulf genetic subpopulation. To allow for the possibility that
different subpopulations had different strengths of connection between
our phenotypes and genotypes (41), we conducted three sets of genetic
analyses: on Gulf and Midwest genotypes separately, and on both
subpopulations together (`Both' subpopulations). Analyses to determine
narrow-sense heritability (h\textsuperscript{2}) for green-up and
flowering were done using linear mixed models and followed (33). Details
on these models can be found in (SI Appendix, Section S6).

@@

To evaluate the prevalence and kinds of covariance patterns of SNP
effects across our common gardens, we used multivariate adaptive
shrinkage (\emph{mash}) on SNP effect estimates from the diversity panel
(36). To do this, we first conducted univarate genome-wide association
(GWAS) on site-specific best linear unbiased predictors (BLUPs) for the
timing of vegetative growth and flowering, using the switchgrassGWAS
package and the methods in (40) (SI Appendix, Section S2). We used the
effect estimates for single nucleotide polymorphisms (SNPs) from these
GWAS as preliminary effect estimates for the two traits at each garden.
We then jointly modeled genetic effect estimates, and the mixture of
ways these effects might covary, across all eight common gardens, using
a two-part procedure: a greedy algorithm to choose covariance structures
using a random SNP subset, followed by a final run of mash with the best
covariance structures using a strong SNP subset (SI Appendix, Section
S3-S5). In brief, we first modeled genetic effects with different
combinations of covariance structures, using the effect size estimates
from subsets of 71-127K relatively unlinked (r\textsuperscript{2}
\textless{} 0.2), randomly selected SNPs used in the GWAS. The number of
SNPs was chosen to result in an approximately million entry matrix for
each trait \& subpopulation, where matrix rows were SNPs and columns
were univariate GWAS. In these models (hereafter referred to as
``random''), we then specified a variety of potential covariance
structures among genetic effect size estimates from each garden (as
expanded on below) and used a greedy algorithm to iteratively select and
add covariance structures that significantly improved the log likelihood
of the \emph{mash} model. Given the high correlation between some of the
covariance patterns we tested, this approach allowed us to select only
the subset of patterns of covariance that significantly improved the fit
of the jointly inferred structure of genetic effect estimates present in
the data. We then selected 19K - 33K relatively unlinked
(r\textsuperscript{2} \textless{} 0.2) SNPs from each univariate GWAS
that had the largest effect estimates in each GWAS (hereafter ``strong''
effects). Notably, the majority of these ``strong'' effects were not
significant in univariate GWAS. In many cases, there was overlap in the
markers with the largest effect estimates at different locations; thus,
19 - 33K SNPs selected for each of eight univariate GWAS ultimately gave
sets of 35K - 63K SNPs. We jointly re-estimated these effect size
estimates at all eight sites using the set of significant covariance
matrices that most improved the model fit on the random effects (SI
Appendix, Section S5, Datasets 1-6).

We generated hypothesis-based covariance matrices derived from
correlations in environmental cues in the green-up or flowering date
windows for the three populations (Table 1; SI Appendix, Section S1).
These covariance matrices represent correlations between identical
genotypes drawn from a specific population at pairs of common gardens;
covariances near one mean that the population has a strong, positive
linear relationship in individual responses at that pair of gardens,
while covariances near zero mean that there is no relationship within
the population for individual responses at that pair of gardens.
\emph{Mash} SNP effects will undergo strong shrinkage towards one
another in the first case, and little shrinkage in the second case.
\emph{Mash} also generates data-driven covariance matrices corresponding
to major patterns of SNP effects present in the data. We generated six
data-driven matrices per \emph{mash} run, five produced by singular
value decomposition (SVD) of an overall matrix.

@@

To confirm candidate genomic regions and patterns of allelic effects
found in the diversity panel, we analyzed flowering in an outbred
pseudo-F2 cross between four individuals, two Midwest and two Gulf
individuals. The formation of this mapping population has been described
previously (35); additional details on QTL mapping can be found in SI
Appendix, Section S7. To be directly comparable to the diversity panel
data, only 2019 phenology data from the pseudo-F2 cross from the same
eight common garden sites were used.
